---
title: "Randall_Plyler_CH3-4 EOC"
author: "Randall Plyler"
date: "1/22/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Sales of Riding Mowers: Scatter Plots. A company that manufactures riding
mowers wants to identify the best sales prospects for an intensive sales campaign. In
particular, the manufacturer is interested in classifying households as prospective owners or nonowners on the basis of Income (in $1000s) and Lot Size (in 1000 ft2). 

The marketing expert looked at a random sample of 24 households, given in the file RidingMowers.csv.
a. Using R, create a scatter plot of Lot Size vs. Income, color-coded by the outcome
variable owner/nonowner. Make sure to obtain a well-formatted plot (create legible labels and a legend, etc.).


```{r 3.2}
library(ggplot2)

df <- read.csv("C:/Users/randa/Dropbox/Masters/Winter/TBANLT 560 Data Mining/Files/DMBA-R-datasets/DMBA-R-datasets/RidingMowers.csv", header = TRUE)
show(df)
ggplot(df, aes(x=Lot_Size, y=Income, colour=Ownership)) + geom_point(shape=19, size=2.5)

```



3 Laptop Sales at a London Computer Chain: Bar Charts and Boxplots. The
file LaptopSalesJanuary2008.csv contains data for all sales of laptops at a computer chain
in London in January 2008. This is a subset of the full dataset that includes data for
the entire year.

a. Create a bar chart, showing the average retail price by store. Which store has the
highest average? Which has the lowest?

b. To better compare retail prices across stores, create side-by-side boxplots of retail
price by store. Now compare the prices in the two stores from (a). Does there
seem to be a difference between their price distributions?


```{r 3.3A}
library(dplyr)
library(ggplot2)

london_data <- read.csv("C:/Users/randa/Dropbox/Masters/Winter/TBANLT 560 Data Mining/Files/DMBA-R-datasets/DMBA-R-datasets/LaptopSalesJanuary2008.csv", header = TRUE)  # load data
london_data2 <- london_data
#Store.Postcode
#Retail.Price

#london_data <- select(london_data_original, c('Store.Postcode', 'Retail.Price'))
london_data <- aggregate(london_data$Retail.Price, by = list(london_data$Store.Postcode), FUN = mean)
barplot(london_data$x, names.arg =london_data$Group.1, xlab = "Location", ylab = "Average Price", las=2)

```
```{r 3.3B}

ggplot(london_data2)+ geom_boxplot(aes(london_data2$Store.Postcode,london_data2$Retail.Price))+theme(axis.text.x = element_text(angle = 90))

```

```
```
Breakfast Cereals. Use the data for the breakfast cereals example in Section 4.8 to
explore and summarize the data as follows:
a. Which variables are quantitative/numerical? Which are ordinal? Which are
nominal?
b. Compute the mean, median, min, max, and standard deviation for each of the
quantitative variables. This can be done through R’s sapply() function (e.g., sapply(data, mean, na.rm = TRUE)).
c. Use R to plot a histogram for each of the quantitative variables. Based on the
histograms and summary statistics, answer the following questions:
i. Which variables have the largest variability?
ii. Which variables seem skewed?
iii. Are there any values that seem extreme?
d. Use R to plot a side-by-side boxplot comparing the calories in hot vs. cold cereals.
What does this plot show us?
e. Use R to plot a side-by-side boxplot of consumer rating as a function of the shelf
height. If we were to predict consumer rating from shelf height, does it appear that
we need to keep all three categories of shelf height?
f. Compute the correlation table for the quantitative variable (function cor()). In addition, generate a matrix plot for these variables (function plot(data)).
i. Which pair of variables is most strongly correlated?
ii. How can we reduce the number of variables based on these correlations?
iii. How would the correlations change if we normalized the data first?
g. Consider the first PC of the analysis of the 13 numerical variables in Table 4.11.
Describe briefly what this PC represents.

```{r 4 a b}

cerealData <- read.csv("C:/Users/randa/Dropbox/Masters/Winter/TBANLT 560 Data Mining/Files/DMBA-R-datasets/DMBA-R-datasets/cereals.csv", header = TRUE)
summary(cerealData) #calculate the summary statistics of the variables

#name ->nominal 
#mfr ->nominal  
#type ->nominal
#calories ->numerical
#protein ->numerical
#fat ->numerical
#sodium ->numerical
#fiber ->numerical
#carbo ->numerical
#sugars ->numerical
#potass ->numerical
#vitamins ->numerical
#shelf ->ordinal
#weight ->numerical
#cups ->numerical
#rating ->ordinal

#This text is from the book and what I used to classify the variables
#Categorical variables can be either coded as numerical (1, 2, 3) or text

#(payments current, payments not current, bankrupt). Categorical variables can
#be unordered (called nominal variables) with categories such as North America,
#Europe, and Asia; or they can be ordered (called ordinal variables) with categories
#such as high value, low value, and nil value.
```
```{r 4 b}
standDev <- sapply(cerealData, sd)
standDev 
#This shows all of the standard deviation's of all the attributes that have them.
```
```{r c}
library(ggplot2)
#name ->nominal 
#mfr ->nominal  
#type ->nominal
#calories ->numerical
#protein ->numerical
#fat ->numerical
#sodium ->numerical
#fiber ->numerical
#carbo ->numerical
#sugars ->numerical
#potass ->numerical
#vitamins ->numerical
#shelf ->ordinal
#weight ->numerical
#cups ->numerical
#rating ->ordinal
ggplot(cerealData, aes(x=calories)) + geom_histogram()
ggplot(cerealData, aes(x=protein)) + geom_histogram()
ggplot(cerealData, aes(x=fat)) + geom_histogram()
ggplot(cerealData, aes(x=sodium)) + geom_histogram()
ggplot(cerealData, aes(x=fiber)) + geom_histogram()
ggplot(cerealData, aes(x=carbo)) + geom_histogram()
ggplot(cerealData, aes(x=sugars)) + geom_histogram()
ggplot(cerealData, aes(x=potass)) + geom_histogram()
ggplot(cerealData, aes(x=vitamins)) + geom_histogram()
ggplot(cerealData, aes(x=shelf)) + geom_histogram()
ggplot(cerealData, aes(x=weight)) + geom_histogram()
ggplot(cerealData, aes(x=cups)) + geom_histogram()
ggplot(cerealData, aes(x=rating)) + geom_histogram()

```

```{r d}
boxplot(calories~type,data=cerealData, xlab="Hot and Cold cereals",ylab="Calories")
```

```{r e}
boxplot(rating~shelf,data=cerealData,xlab="rating",ylab="shelf",horizontal=TRUE)

```

```{r f}
#name ->nominal 
#mfr ->nominal  
#type ->nominal
#calories ->numerical
#protein ->numerical
#fat ->numerical
#sodium ->numerical
#fiber ->numerical
#carbo ->numerical
#sugars ->numerical
#potass ->numerical
#vitamins ->numerical
#shelf ->ordinal
#weight ->numerical
#cups ->numerical
#rating ->ordinal

#This is a correlation matrix of the numberic and ordinal variables.
dataframeforcorrelationmatrix <- cerealData %>% select(calories,protein,fat,sodium,fiber,carbo,sugars,potass,vitamins,shelf,weight,cups,rating)
show(dataframeforcorrelationmatrix)
correlation_matrix2 <- cor(dataframeforcorrelationmatrix)
correlation_matrix2

#i. Which pair of variables is most strongly correlated?
#calories and weight with a .6960 correlation
#ii. How can we reduce the number of variables based on these correlations?
#by removing the highly correlated variables the collinearity will improve
#iii. How would the correlations change if we normalized the data first?
#The correlation will not change
```

